Exploring the Design Space of Differentially Private Bayesian Inference under Hellinger Distance

### 2019-01-23 ICML2019/main.tex
tweak

%  support bayesian inference
% by means of a language primitive, often called $\bysinfer$, which
% supports the (precise or approximate) parametric computation of the
% posterior of a given model. In the rest of the paper, we will use the following
% notation as a shorthand for the inference process:
% \[
%   \bysinfer(\Pr(\dataobs | \theta), \Pr(\theta), \dataobs) = \frac{\Pr(\dataobs | \theta) \cdot \Pr(\theta)}{\Pr(\dataobs)}
% \]
\noindent \textbf{Dirichlet-multinomial model.}
In this work we will focus on the Dirichlet-Multinomial model for
categorical data and its restriction to binary data: the Beta-Binomial
model. 
That is, we will consider the situation where 
the underlying categorical data is drawn from a multinomial
distribution with parameter the vector $\vtheta\in [0,1]^{k}$. The prior distribution over $\vtheta$
is given by a Dirichlet distribution, $\dirichlet{\valpha}$, for,
$\valpha\in(\mathbb{R}^{+})^{k}$ and $k\in\mathbb{N}$, with p.d.f:
\[
\Pr(\vtheta)=\frac{1}{\mbetaf(\valpha)}\cdot\prod_{i=1}^{k}{\theta_i^{\alpha_i-1}}
\]
where $\mbetaf(\cdot)$ is the generalized beta function.
The data $\dataobs$ is a sequence of $n\in\mathbb{N}$ values
coming from a universe $\datauni$, such that $|\datauni|=k$.
The likelihood function is
$
\Pr(\dataobs|\vtheta)=\prod_{a_i\in\datauni}\theta_{i}^{\Delta \alpha_i},
$
with $\Delta \alpha_i=\displaystyle\sum_{j=1}^{n}\mathbf{1}_{\iverson{x_j=a_i}}$.
Denoting by $\Delta\valpha$ the vector $(\Delta\alpha_1,\dots
\Delta\alpha_k)$ the posterior distribution over $\vtheta$ is
$\dirichlet{\valpha + \Delta \valpha}$, where $+$ denotes the componentwise sum of vectors.
In the following we will use the notation $\bysdir(n,\vtheta, \valpha,
\dataobs)$ to denote the posterior of this process.
%i.e., $\bysdir(n,\vtheta, \valpha, \dataobs)=\dirichlet{\valpha + \Delta \valpha}$.
%
% ($\thicksim \binomial{n}{\theta}$), where $\theta$ represents
% the parameter --informally called \emph{bias}-- of a Bernoulli
% distributed random variable and $n$ is the number of draws. The
% prior distribution over $\theta\in [0,1]$ is going to be a beta
% distribution, $\betad(\alpha, \beta)$, with parameters
% $\alpha,\beta\in\mathbb{R}^{+}$, and with p.d.f:
% \[
%   \Pr(\theta)= \frac{\theta^{\alpha} (1- \theta)^{\beta}}{\betaf(\alpha,\beta)}
% \]
% where $\betaf(\cdot,\cdot)$ is the beta function.
% The data $\dataobs$ will be a sequence of $n\in\mathbb{N}$ binary values, that is $\dataobs= (x_1,\dots x_n), x_i\in\{0,1\}$, and the likelihood function is:
% \[
%   \Pr(\dataobs | \theta)= \theta^{\Delta \alpha}(1-\theta)^{n - \Delta \alpha}
% \]
% where $\Delta \alpha = \sum_{i=1}^{n}x_i$.
% The posterior distribution over $\theta$ can be easily proven to be equal to:\[\betad(\alpha + \Delta \alpha,\beta + n - \Delta \alpha)\]
% From now on we will denote by $\bysinfer(\binomial{n}{\theta}, \betad(\alpha, \beta), \dataobs)$ the previous
% posterior distribution. This notation will prove itself handy when its arguments will change types in the next example and
% in section \ref{} when the private mechanisms will be described.

@@

\subsection{Calibrating noise w.r.t. $\ell_1$ norm ($\lapmech$)}
\label{sec_ell1_dim}
We start off by presenting $\lapmech$, in Algorithm \ref{lapmech}, based on the Laplace primitive.
The idea of $\lapmech$ is to add Laplacian noise with scale proportional to $\mathcal{X}$ to
the differences between the respective numeric parameters of the prior
and posterior distribution. 
\begin{algorithm}
  \caption{$\lapmech$}
  \label{lapmech}
  \begin{algorithmic}
  \INPUT $\dataobs\in\mathcal{X}^n$, $\dirichlet{\valpha}$
  \STATE \quad {\bf let} $\valpha' = \inferd{\valpha}{\vtheta}{\dataobs}$
  \STATE \quad {\bf Initialize} a vector $\tilde{\valpha} = (0, \dots, 0 )\in \mathbb{N}^{|\mathcal{X}|}$ 
  \STATE \quad {\bf For} $i = 1 \dots |\mathcal{X}| - 1$:
  \STATE \quad \quad  {\bf let} $\eta \sim \lap{0}{\frac{|\mathcal{X}|}{\epsilon}}$
  \STATE \quad \quad  $\tilde{\alpha_i}=\alpha_i + \lfloor{(\alpha_i' - \alpha_i) + \eta}\rfloor^n_0$ 
  % \RETURN $a_{k+1}$.
  \STATE \quad $\tilde{\alpha_{|\mathcal{X}|}} = \alpha_{|\mathcal{X}|} + \lfloor n - \sum_{i = 1}^{|\mathcal{X}|-1}\lfloor{(\alpha_i' - \alpha_i) + \eta_i}\rfloor^n_0 \rfloor^n_0$
  \STATE {\bf return} $\tilde{\valpha}$
  \end{algorithmic}
\end{algorithm}
\begin{lem}
  Algorithm \ref{lapmech} is $\epsilon$-d.p.
  \begin{proof} By noticing that $\mathcal{X}-1$ is an upper bound to the value denoted by (\ref{eqn:eq1}). \end{proof}
\end{lem}

>>>>

%  support bayesian inference
% by means of a language primitive, often called $\bysinfer$, which
% supports the (precise or approximate) parametric computation of the
% posterior of a given model. In the rest of the paper, we will use the following
% notation as a shorthand for the inference process:
% \[
%   \bysinfer(\Pr(\dataobs | \theta), \Pr(\theta), \dataobs) = \frac{\Pr(\dataobs | \theta) \cdot \Pr(\theta)}{\Pr(\dataobs)}
% \]
\noindent \textbf{Dirichlet-multinomial model.}
In this work we will focus on the Dirichlet-Multinomial model for
categorical data and its restriction to binary data: the Beta-Binomial
model. 
That is, we will consider the situation where 
the underlying categorical data is drawn from a multinomial
distribution with parameter the vector $\vtheta\in [0,1]^{k}$. The prior distribution over $\vtheta$
is given by a Dirichlet distribution, $\dirichlet{\valpha}$, for,
$\valpha\in(\mathbb{R}^{+})^{k}$ and $k\in\mathbb{N}$, with p.d.f:
\[
\Pr(\vtheta)=\frac{1}{\mbetaf(\valpha)}\cdot\prod_{i=1}^{k}{\theta_i^{\alpha_i-1}}
\]
where $\mbetaf(\cdot)$ is the generalized beta function.
The data $\dataobs$ is a sequence of $n\in\mathbb{N}$ values
coming from a universe $\datauni$, such that $|\datauni|=k$.
The likelihood function is
$
\Pr(\dataobs|\vtheta)=\prod_{a_i\in\datauni}\theta_{i}^{\Delta \alpha_i},
$
with $\Delta \alpha_i=\displaystyle\sum_{j=1}^{n}\mathbf{1}_{\iverson{x_j=a_i}}$.
Denoting by $\Delta\valpha$ the vector $(\Delta\alpha_1,\dots
\Delta\alpha_k)$ the posterior distribution over $\vtheta$ is
$\dirichlet{\valpha + \Delta \valpha}$, where $+$ denotes the componentwise sum of vectors.
In the following we will use the notation $\bysinfer(\valpha,\vtheta,
\dataobs)$ to denote the posterior of this process.
%i.e., $\bysdir(n,\vtheta, \valpha, \dataobs)=\dirichlet{\valpha + \Delta \valpha}$.
%
% ($\thicksim \binomial{n}{\theta}$), where $\theta$ represents
% the parameter --informally called \emph{bias}-- of a Bernoulli
% distributed random variable and $n$ is the number of draws. The
% prior distribution over $\theta\in [0,1]$ is going to be a beta
% distribution, $\betad(\alpha, \beta)$, with parameters
% $\alpha,\beta\in\mathbb{R}^{+}$, and with p.d.f:
% \[
%   \Pr(\theta)= \frac{\theta^{\alpha} (1- \theta)^{\beta}}{\betaf(\alpha,\beta)}
% \]
% where $\betaf(\cdot,\cdot)$ is the beta function.
% The data $\dataobs$ will be a sequence of $n\in\mathbb{N}$ binary values, that is $\dataobs= (x_1,\dots x_n), x_i\in\{0,1\}$, and the likelihood function is:
% \[
%   \Pr(\dataobs | \theta)= \theta^{\Delta \alpha}(1-\theta)^{n - \Delta \alpha}
% \]
% where $\Delta \alpha = \sum_{i=1}^{n}x_i$.
% The posterior distribution over $\theta$ can be easily proven to be equal to:\[\betad(\alpha + \Delta \alpha,\beta + n - \Delta \alpha)\]
% From now on we will denote by $\bysinfer(\binomial{n}{\theta}, \betad(\alpha, \beta), \dataobs)$ the previous
% posterior distribution. This notation will prove itself handy when its arguments will change types in the next example and
% in section \ref{} when the private mechanisms will be described.

@@

\subsection{Calibrating noise w.r.t. $\ell_1$ norm ($\lapmech$)}
\label{sec_ell1_dim}
We start off by presenting $\lapmech$, in Algorithm \ref{lapmech}, based on the Laplace primitive.
The idea of $\lapmech$ is to add Laplacian noise with scale
proportional to $\mathcal{X}$ (the dimension of the parameter space) to
the differences between the respective numeric parameters of the prior
and posterior distribution. 
\begin{algorithm}
  \caption{$\lapmech$}
  \label{lapmech}
  \begin{algorithmic}
  \INPUT $\dataobs\in\mathcal{X}^n$, $\dirichlet{\valpha}$
  \STATE \quad {\bf let} $\valpha' = \inferd{\valpha}{\vtheta}{\dataobs}$
  \STATE \quad {\bf Initialize} a vector $\tilde{\valpha} = (0, \dots, 0 )\in \mathbb{N}^{|\mathcal{X}|}$ 
  \STATE \quad {\bf For} $i = 1 \dots |\mathcal{X}| - 1$:
  \STATE \quad \quad  {\bf let} $\eta \sim \lap{0}{\frac{|\mathcal{X}|}{\epsilon}}$
  \STATE \quad \quad  $\tilde{\alpha_i}=\alpha_i + \lfloor{(\alpha_i' - \alpha_i) + \eta}\rfloor^n_0$ 
  % \RETURN $a_{k+1}$.
  \STATE \quad $\tilde{\alpha_{|\mathcal{X}|}} = \alpha_{|\mathcal{X}|} + \lfloor n - \sum_{i = 1}^{|\mathcal{X}|-1}\lfloor{(\alpha_i' - \alpha_i) + \eta_i}\rfloor^n_0 \rfloor^n_0$
  \STATE {\bf return} $\tilde{\valpha}$
  \end{algorithmic}
\end{algorithm}
\begin{lem}
  Algorithm \ref{lapmech} is $\epsilon$-d.p.
  \begin{proof} By noticing that $\mathcal{X}-1$ is an upper bound to the value denoted by (\ref{eqn:eq1}). \end{proof}
\end{lem}
